Unsupervised embedding of trajectories captures the latent structure of mobility

### 2020-10-01 paper/current/main.tex
abstract edits

\begin{document} %{{{
\title{Learning a functional distance between organizations with unsupervised embeddings of scientific mobility}
\title{Unsupervised embeddings of trajectories captures the latent structure of mobility} %{{{
\date{\today}
\maketitle %}}}

@@

Human mobility and migration drive major societal phenomena such as the growth and evolution of cities~\autocite{wef2017migration, curiel2018citygravity},  epidemics~\autocite{kraemer2020covid, truscott2012epidemicgravity, xia2005measlesgravity}, economies~\autocite{kaluza2010cargo, kerr2011immigration}, and innovation~\autocite{kaiser2018innovation, armano2017innovation, sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage}.
Historically, human mobility is strongly constrained by physical separation---geographical distance~\autocite{zipf1946gravity}.
However, geographic distance is becoming less relevant in the increasingly-globalized world~\autocite{czaika2018migration, czaika2018globalisation, buchan2009globalization, altbach2004globalization} in which physical barriers are shrinking while linguistic, cultural, and historical relationships are becoming more important, which calls for an alternative modeling framework for human mobility that takes these factors into account~\autocite{boschma2005proximity,  brown1970functional, brown1970migration, kim2018functional}.
Here, using a massive dataset of scientific mobility and U.S. flight itineraries, we demonstrate that a neural embedding model can successfully encode nuanced relationships between locations into a vector-space ~\autocite{mikolov2013word2vec}, providing an effective functional distance that reflects the multi-faced structure of human mobility. 
Focusing on the case of scientific mobility, we show that the embedding captures cultural and linguistic relations at multiple levels of granularity, and even academic prestige.
Furthermore, embedding vectors themselves also reveal insights on the role of the organization in the landscape of global scientific mobility universally.
The ability to learn scalable, dense and meaningful high-resolution vector representations of mobility directly from the data can open up a new avenue of studying mobility across domains.

>>>>

\begin{document} %{{{
\title{Unsupervised embedding of trajectories captures the latent structure of mobility} %{{{
\date{\today}
\maketitle %}}}

@@

Human mobility and migration drive major societal phenomena such as the growth and evolution of cities~\autocite{wef2017migration, curiel2018citygravity},  epidemics~\autocite{kraemer2020covid, truscott2012epidemicgravity, xia2005measlesgravity}, economies~\autocite{kaluza2010cargo, kerr2011immigration}, and innovation~\autocite{kaiser2018innovation, armano2017innovation, sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage}.
Historically, human mobility is strongly constrained by physical separation---geographical distance~\autocite{zipf1946gravity}.
However, geographic distance is becoming less relevant in the increasingly-globalized world~\autocite{czaika2018migration, czaika2018globalisation, buchan2009globalization, altbach2004globalization} in which physical barriers are shrinking while linguistic, cultural, and historical relationships are becoming more important, which calls for an alternative modeling framework for human mobility that takes these factors into account~\autocite{boschma2005proximity,  brown1970functional, brown1970migration, kim2018functional}.
Here, using a massive dataset of scientific mobility and U.S. flight itineraries, we demonstrate that a neural embedding model can successfully encode nuanced relationships between locations into a vector-space~\autocite{mikolov2013word2vec}, providing an effective functional distance that reflects the multi-faced structure of human mobility. 
Focusing on the case of scientific mobility, we show that embeddings of scientific organizations captures cultural and linguistic relations at multiple levels of granularity, and even academic prestige.
Furthermore, the embedding vectors themselves reveal universal relationships between organizational characteristics, and their place in the global landscape of scientific mobility. 
The ability to learn scalable, dense and meaningful high-resolution vector representations of mobility directly from the data can open up a new avenue of studying mobility across domains.
